Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-04 paper/paper.tex
pass on 2.2

Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}.
A Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The action part appears as the first
word in the rule description. An action denotes the task that needs to
be executed if the rule pattern is satisfied.  In this example, a
message will be sent to system administrators if the rule pattern is
satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
  differs from an IDS in the action taken when an intrusion is
  observed. IPS systems block malicous traffic. However, IDS and IPS
  do not differ in the approaches that can be used to detect the
  intrusion~\cite{ids-ips}.\Mar{check}}. The header comes after the action in the rule
description. It restricts the information flow covered by the
rule. The header of this rule is \CodeIn{icmp any any -> any
  any}. It instructs Suricata to inspect \CodeIn{icmp} traffic flowing
from any machine and port to any machine and port.\MyComment{ The variables
\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
configurable.} The rule options come after the header. It is a
semi-colon-separated sequence of key-value pairs. Rule options serve
to document the analyzed traffic (\eg{}, ``msg'', ``classtype'', and
``sid'') and to characterize the attack pattern (\eg, ``flags'',
``threshold'', and ``content''). \tname{} produces options of the
second kind, which are \emph{relevant} to capture the attack. For
example, the option \CodeIn{content:<arg>} checks if the argument is
present in the payload of the message.

>>>>

Figure~\ref{fig:pingscan-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}.  A
Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The \emph{action} appears as the
first word in the rule description (an \CodeIn{alert}, in this
case). An action denotes the task that needs to be executed if the
rule pattern is satisfied.  In this example, an alert message will be
sent to system administrators if the rule pattern is
satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
  differs from an IDS in the action taken when an intrusion is
  observed. IPS systems block malicous traffic whereas IDS system only
  alert them. However, IDS and IPS do not differ in the approaches
  used to detect the intrusion~\cite{ids-ips}.\Mar{check}}. The
\emph{header} comes after the action in the rule description. It
restricts the information flow covered by the rule. The header of this
rule is \CodeIn{icmp any any -> any any}. It instructs Suricata to
inspect \CodeIn{icmp} traffic flowing from any machine and port to any
machine and port.\MyComment{ The variables \CodeIn{\$HOME\_NET} and
  \CodeIn{\$EXTERNAL\_NET} are configurable.} Finally, \emph{rule
  options} come after the header. It is a semi-colon-separated
sequence of key-value pairs. Rule options serve to document the
analyzed traffic (\eg{}, ``msg'', ``classtype'', and ``sid'') and to
characterize the attack pattern (\eg, ``flags'', ``threshold'', and
``content''). \tname{} produces options of the second kind, which are
\emph{relevant} to capture the attack. For example, the option
\CodeIn{content:<arg>} checks if the argument is present in the
payload of the message.
